\bf The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-08-06 paper/paper.tex
results for appedices 3+4 need to edit text

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
\CommentTok{## # A tibble: 6,296 x 5}
\CommentTok{##   focal_ID focal_sp   geometry growth comp             }
\CommentTok{##      <dbl> <fct>       <POINT>  <dbl> <list>           }
\CommentTok{## 1        4 nysy     (14.2 428)  0.103 <tibble [20 x 4]>}
\CommentTok{## 2        5 havi      (9.4 436)  0.150 <tibble [32 x 4]>}
\CommentTok{## 3       79 tiam       (40 381) -0.161 <tibble [20 x 4]>}
\CommentTok{## 4       80 caca     (38.7 422)  0.253 <tibble [12 x 4]>}
\CommentTok{## 5       96 libe       (60 310)  0.262 <tibble [14 x 4]>}
\CommentTok{## # ... with 6,291 more rows}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
\CommentTok{## # A tibble: 20 x 6}
\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_basal_area}
\CommentTok{##      <dbl> <dbl>   <dbl> <dbl> <fct>             <dbl>}
\CommentTok{## 1        4  13.6    1836  7.48 tiam            0.0176 }
\CommentTok{## 2        4  13.6    1847  2.81 nysy            0.00332}
\CommentTok{## 3        4  13.6    1848  1.62 nysy            0.00396}
\CommentTok{## 4        4  13.6    1849  2.62 nysy            0.00535}
\CommentTok{## 5        4  13.6    1850  2.98 havi            0.00472}
\CommentTok{## # ... with 15 more rows}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}

@@

@@

\CommentTok{# Compute focal versus competitor tree information ---------------------}
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}

@@

  \CommentTok{# Make predictions and compute RMSE}
\NormalTok{  shuffle_RMSE[j] <-}\StringTok{  }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} 
             \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
\NormalTok{\}}

@@

\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}

@@

\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# make all species the same, needs to be factor with at least two levels}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\StringTok{'B'}\NormalTok{)) )}

@@

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist))}
\NormalTok{\{}
\NormalTok{  comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}

@@

\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}

@@

\begin{center}\includegraphics[width=1\linewidth]{Figures/appendix2-code-1} \end{center}

@@

>>>>

We also load species information as it existed on GitHub on 2021/08/02,
which includes family, genus, and species information; as well as
classifications of the canopy position (canopy, canopy emergent,
understory, shrub layer) and drought tolerance (intolerant, resistant)
of the species.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sp_info <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"SCBI_ForestGEO_sp_ecology.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}
    \DataTypeTok{sp =}\NormalTok{ spcode, family, genus, species, canopy_position, }
\NormalTok{    drought_tolerance}
\NormalTok{  ) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{()}
\NormalTok{sp_info}
\CommentTok{## # A tibble: 65 x 6}
\CommentTok{##   sp    family        genus       species     canopy_position drought_toleran~}
\CommentTok{##   <chr> <chr>         <chr>       <chr>       <chr>           <chr>           }
\CommentTok{## 1 acne  Sapindaceae   Acer        negundo     understory      drought-resista~}
\CommentTok{## 2 acpl  Sapindaceae   Acer        platanoides canopy          drought-resista~}
\CommentTok{## 3 acru  Sapindaceae   Acer        rubrum      canopy          drought-resista~}
\CommentTok{## 4 aial  Simaroubaceae Ailanthus   altissima   canopy          drought-resista~}
\CommentTok{## 5 amar  Rosaceae      Amelanchier arborea     understory      drought-resista~}
\CommentTok{## # ... with 60 more rows}
\end{Highlighting}
\end{Shaded}

We join this species information to our \texttt{growth\_scbi} data
frame.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\CommentTok{# }\AlertTok{TODO}\CommentTok{:}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
\end{Highlighting}
\end{Shaded}

Furthermore, we compute two potential competitor explanatory variables
\(x^{\text{comp}}_{ijk}\) from Equation \ref{eq:model}. First, the basal
area of each tree as a function of its DBH in the earlier census.
Second, the above ground biomass as estimated by allometric equations
encoded in the \texttt{get\_biomass()} function from the \texttt{allodb}
package \citep{allodb_package}; this function has DBH, species, and
geographic coordinates as arguments.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Install development version of allodb using:}
\CommentTok{# remotes::install_github("forestgeo/allodb")}
\KeywordTok{library}\NormalTok{(allodb)}
\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \CommentTok{# Compute basal area:}
    \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{,}
    \CommentTok{# Compute above ground biomass:}
    \DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}
      \DataTypeTok{dbh =}\NormalTok{ dbh1,}
      \DataTypeTok{genus =}\NormalTok{ genus,}
      \DataTypeTok{species =}\NormalTok{ species,}
      \DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)}
\NormalTok{    )}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

@@

\CommentTok{# Associate each observation to a fold}
\NormalTok{spatial_block_scbi <-}\StringTok{ }
\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}
    \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
    \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
\NormalTok{  )}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)}
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
\CommentTok{## # A tibble: 6,296 x 5}
\CommentTok{##   focal_ID focal_sp   geometry growth comp             }
\CommentTok{##      <dbl> <fct>       <POINT>  <dbl> <list>           }
\CommentTok{## 1        4 nysy     (14.2 428)  0.103 <tibble [20 x 4]>}
\CommentTok{## 2        5 havi      (9.4 436)  0.150 <tibble [32 x 4]>}
\CommentTok{## 3       79 tiam       (40 381) -0.161 <tibble [20 x 4]>}
\CommentTok{## 4       80 caca     (38.7 422)  0.253 <tibble [12 x 4]>}
\CommentTok{## 5       96 libe       (60 310)  0.262 <tibble [14 x 4]>}
\CommentTok{## # ... with 6,291 more rows}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
\CommentTok{## # A tibble: 20 x 6}
\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_x_var}
\CommentTok{##      <dbl> <dbl>   <dbl> <dbl> <fct>        <dbl>}
\CommentTok{## 1        4  13.6    1836  7.48 tiam       0.0176 }
\CommentTok{## 2        4  13.6    1847  2.81 nysy       0.00332}
\CommentTok{## 3        4  13.6    1848  1.62 nysy       0.00396}
\CommentTok{## 4        4  13.6    1849  2.62 nysy       0.00535}
\CommentTok{## 5        4  13.6    1850  2.98 havi       0.00472}
\CommentTok{## # ... with 15 more rows}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }
                              \DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  ) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\CommentTok{# Compute basal area:}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}

@@

\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2),}
  \KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}

@@

\CommentTok{# Compute focal versus competitor tree information ---------------------}
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)}

@@

  \CommentTok{# Make predictions and compute RMSE}
\NormalTok{  shuffle_RMSE[j] <-}\StringTok{  }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}
      \DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)}
\NormalTok{    ) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
\NormalTok{\}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}

@@

\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# make all species the same, needs to be factor with at least two levels}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\StringTok{'B'}\NormalTok{)) ) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\CommentTok{# Compute basal area:}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}

@@

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist))\{}
\NormalTok{  comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}

@@

\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}

@@

@@

\hypertarget{compare-pred-var}{%
\section{Appendix 3: Compare predictor
variables}\label{compare-pred-var}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}

\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}

\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp),}
        \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{,}
         \DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}
                  \DataTypeTok{dbh =}\NormalTok{ dbh1,}
                  \DataTypeTok{genus =}\NormalTok{ genus,}
                  \DataTypeTok{species =}\NormalTok{ species,}
                  \DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)}
\NormalTok{    ))}

\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}

\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}

\NormalTok{focal_vs_comp_ba <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
                         \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}

\NormalTok{focal_vs_comp_agb <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
                         \DataTypeTok{comp_x_var =} \StringTok{"agb"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}

\NormalTok{  focal_vs_comp_ba }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.14}

\NormalTok{  focal_vs_comp_agb }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 1.83}
\end{Highlighting}
\end{Shaded}

\hypertarget{compare-group-var}{%
\section{Appendix 4: Compare grouping
variables}\label{compare-group-var}}

Creative use of the grouping variable allows for tests of what best
explains competitive interactions. Here we compare whether species
idenity or canopy position of the species best explains.

\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
